Climate variability indices for ecological and crop models in R: the `climatrends` package

### 2020-08-14 inst/paper/paper.Rmd
paper

>Kauê de Sousa^1,2[*]^, Jacob van Etten^2^, Svein Øivind Solberg^1^  
^1^ Department of Agricultural Sciences, Inland Norway University of Applied Sciences, 2318 Hamar, Norway  
^2^ Bioversity International, 00054 Maccarese, Rome, Italy  
^[*]^Correspondence should be addressed to: <kaue.desousa@inn.no>

@@

## Some GDD related analysis

Use the data from seed germination or crop growth to compute GDD. Maybe a GDD analysis in Scandinavia with data from the reindeer project.

@@

## Time series

Pick some random points in Norway (or Scandinavia??) and check how the trends on temperature indices over the last 20 years.

@@

This work was supported by The Nordic Joint Committee for Agricultural and Food Research (grant num. 202100-2817).

>>>>

>Kauê de Sousa^1,2[*]^, Magne Neby^3^, Jacob van Etten^2^, Svein Øivind Solberg^1^  
^1^ Department of Agricultural Sciences, Inland Norway University of Applied Sciences, 2318 Hamar, Norway  
^2^ Bioversity International, 00054 Maccarese, Rome, Italy  
^3^ Department of Forest and Field Sciences, Inland Norway University of Applied Sciences, 2480 Koppang, Norway    
^[*]^ Correspondence should be addressed to: <kaue.desousa@inn.no>

@@

@@

## Trends in climate variability in Norway and Sweden

Pick some random points in Norway and Sweden and check how the trends on temperature indices over the last 20 years.

## Impact of early thaw on vole populations

From [Johnsen et al (2016)](https://doi.org/10.1002/ece3.2635) and [Boonstra et al (2016)](http://bioscience.oxfordjournals.org/content/early/2016/07/19/biosci.biw080.full.pdf). The snowpack of European boreal forest contains relatively dense snow because of snow falling at temperatures just below freezing and because of the formation of ice crusts owing to thaw–freeze cycles during the relatively mild winters. The formation of ice due to freeze–thaw cycles during winter may be limiting vole availability to bilberries and other edible food items by encrusting them in ice and therefore limiting the peak phase of the cycles. If there falls snow in autumn but then it becomes above zero and the snow starts melting but then it starts freezing again, an ice layer is formed which has important implications in ecology. This might be calculated / predicted with the `late_frost()` function, but then applied as 'early thaw' function.

@@

This work was supported by The Nordic Joint Committee for Agricultural and Food Research (grant num. 202100-2817). We thank Julian Ramirez-Villegas and Marcel Schrijvers-Gonlag for the useful insights and discussion that helped in the development of this study.
